### 2019-08-19 presentation/presentation.tex
template to use

\documentclass{beamer}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}

\usetheme{Warsaw}

\title{Presentación}
\author{Miguel A. Gomez}
\institute{Fundación Universitaria Konrad Lorenz}
\date{20-08-2019}

\begin{document}
	\frame{\titlepage}
	\begin{frame}
		\frametitle{Sample frame title}
		This is a text in the first frame. This is a text in the first frame. This is a text in the first frame.
	\end{frame}
	\begin{frame}
		\alert{highlighted} because it's important.
		Please, don't abuse it.

		\begin{block}{Remark}
			Sample text
		\end{block}

		\begin{alertblock}{Important theorem}
			Sample text in red box
		\end{alertblock}

		\begin{examples}
			Sample text in green box. The title of the block is ``Examples".
		\end{examples}
	\end{frame}
	\begin{frame}
	\frametitle{Two-column slide}	
		\begin{columns}

			\column{0.5\textwidth}
			This is a text in first column.
			$$E=mc^2$$
			\begin{itemize}
				\item First item
				\item Second item
			\end{itemize}

			\column{0.5\textwidth}
			This text will be in the second column
			and on a second tought this is a nice looking
			layout in some cases.
		\end{columns}
	\end{frame}
\end{document}